Variance reduction for MCMC methods via martingale representations

### 2020-03-12 aap/Revision/vr-mcmc-revision-new.tex
truncation

\widehat M_{K,n}^N(f) =\frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\left[\sum_{p=0}^{N+n-l}\widehat{a}_{p,k}(X_{l-1}^{x})\phi_{k}(\xi_{l})\right].

@@

>>>>

\widehat M_{K,n}^N(f) = \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\left[\sum_{p=0}^{N+n-l}\widehat{a}_{p,k}(X_{l-1}^{x})\phi_{k}(\xi_{l})\right].

@@

Yet it is impractical to use estimator of the form $\widehat M_{K,n}^N(f)$, since it may be too expensive to compute for $n$ large enough. Instead we introduce the truncated version of this control variate
\tilde M_{K,n}^N(f) = \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\left[\sum_{p=0}^{N+n-l}\widehat{a}_{p,k}(X_{l-1}^{x})\phi_{k}(\xi_{l})\mathbbm{1}\{ |p-k| < n_{\mathrm{trunc}} \}\right]
where $n_{\mathrm{trunc}} \in \nset$ is a hyperparameter. Validity of this approach follows from the following lemma

\begin{lem}
\label{lem:coef_decrease}
Let $(X^x_{p})_{p \in \nset_0}$ be a $V-$geometrically ergodic homogeneous Markov chain with stationary distribution $\pi$ and $f$ be a Borel function with $\|f\|_{V} < \infty$. Then it holds
\begin{equation}
|\bar a_{r,k}(x)| \leq C\rho^r\|f\|_{V} (\PE V^2(\Phi(x,\xi)))^{1/2}
\end{equation}
where $\bar a_{r,k}(x)$ is defined in \eqref{eq:definition-bar-a}.
\end{lem}
\begin{proof}
Note that 
\[
\bar a_{r,k}(x)=\mathsf{E}\left[\phi_k\left(\xi\right)Q_{r}\left(\Phi(x,\xi)\right)\right]
\]
with  $Q_{r}(x)=\mathsf{E}\left[f(X^x_{r})\right], r\in \mathbb{N}$. $V-$ergodicity of $(X^x_{p})_{p \in \nset_0}$ implies that
\[
|Q_{r}(x) - \pi(f)| \leq |\int\limits_{\rset^d} f(y)\left(\P^r(x,dy)-\pi(dy)\right)| \leq C\rho^r V(x) \|f\|_{V}
\]
Thus, since $\PE\phi_k\left(\xi\right) = 0$, we obtain from Cauchy-Schwarz inequality
\begin{eqnarray*}
|\bar a_{r,k}(x)| \leq \left(\PE \phi^2_k\left(\xi\right)\right)^{1/2}\left(\PE |Q_{r}(\Phi(x,\xi)) - \pi(f)|^{2}\right)^{1/2} \leq C\rho^r\|f\|_{V} (\PE V^2(\Phi(x,\xi)))^{1/2}
\end{eqnarray*}
and the statement follows.
\end{proof}
